\LARGE Untangling the link between traits, size and growth rate in plants

### 2016-08-04 ms/ms.tex
Writing introduction and model

@@

% Enable cross referencing to Evidence document
\usepackage{xr}
\externaldocument{ms-supp}
\externaldocument{figure_conceptual}

\input{ms-common-defs}
\date{}

@@

\keywords{growth strategy, life-history traits, maximum height, ontogenetic stage, photosynthetic rate, plant development, seed mass, specific leaf area, wood density}

%appendix
\usepackage[title,titletoc,toc]{appendix}
\usepackage{setspace}  % for double spacing

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
% \doublespacing
% \linenumbers
\section{Summary}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item We ...
\end{enumerate}

% Consider Ecol Letters I deas and Perspectives: maximum of 7500 words (main text), and 10 figures, tables or boxes.

\section{To do}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item New refs: \citep{Ishihara-2016, Duursma-2016, Gibert-2016, Falster-2016, Kunstler-2015,Falster-2015b, Falster-2015a,Wenk-2015,Li-2014}
\item Table near intro summarising empirical patterns to be explained by integrated model?
\end{enumerate}

@@

Functional traits capture core differences in the strategies
plants use to generate and invest surplus energy
 \citep{Wright-2004, Chave-2009, Westoby-2002}.
Although most plants have the same basic physiological function
and key resource requirements (carbon, nitrogen and water), species differ considerably in
rates at which resources are acquired and invested into different tissues.
Much of this functional variation relates to underlying
differences in some prominent traits, including leaf
mass per area (LMA), wood density (WD), leaf nitrogen per area, vessel
diameter,  maximum  height (HMAX) and seed size  \citep{Wright-2004, Chave-2009}.
Data for these traits now exists for up to 10\% of the world's 250000 plant species
\citep{Cornwell-2014}. Increasingly, researchers have been looking to traits
as potential indicators of the growth strategy and ecology of a species.

While the influence of plant traits on elements of plant physiological
function is increasingly understood, attempts at using traits to understand
demographic rates have met with mixed success
 \citep{Wright-2010, Poorter-2008}. In seedlings, the trait LMA
-- the central element of the leaf economics spectrum
 \citep{Wright-2004} -- has been tightly related to relative growth
rate in plant mass \citep{Lambers-1992, Wright-2000}. LMA and
its correlate leaf lifespan have also been linked to height growth rate for
small seedlings and saplings \citep{Reich-1992, Poorter-2006}. These
early successes prompted researchers to search for similar relationships in
large plants. However, the results showed that in trees LMA was not correlated
with diameter growth rate  \citep{Wright-2010, Poorter-2008,
Herault-2011}. Meanwhile, other traits such as wood density showed
strong relationships to growth in large plants \citep{Wright-2010},
but only weak relationships in seedlings  \citep{Castro-1998}. Thus
far, most trait growth relationships have been quantified for one or few
traits and at a single life- stage. Yet the picture is emerging is that the
link between traits and growth may be modified by plant size
 \citep{Ruger-2012}. This finding challenges a wide-spread
assumption regarding traits: that particular ends of the trait spectrum translate directly into
faster or slower growth (refs XX).

The challenge of interpreting diverse empirical results seeking to link
traits to growth rate is made all the harder because we currently lack clear
expectations on the expected signal. Generating such expectations is one of the
primary role for theory \citep{Kokko-2007}. Current empirical
results suggest the effect of traits on growth changes with size and possibly
also light environment, yet current theory says little about how such
relationships might come about. A widely-used model shows that for seedlings,
mass-based relative growth rate is linearly and negatively related to LMA
 \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
 An extension of the model suggests a similar
relationship may hold at larger sizes \citep{Enquist-2007}. But as
described above, this prediction is not supported by empirical results.
Meanwhile, theoretical predictions on how other traits should influence growth
are largely absent.

A further problem with existing theory is that the effects of traits are all
realised via influences on mass production \citep{Enquist-2007},
whereas the effect of traits such as LMA, WD, and Hmax is via allocated
to different tissues \citep{Falster-2011}. There are in fact two problems with theory
focussed solely on mass production. The first is that measuring mass
production is only practical for small plants that can be easily harvested.
For larger plants, growth is more often measured as increment in either diameter
or height. For example, the CTFS monitors diameter growth in 49 large permanent
plots across forests worldwide \citep{Anderson-2015};
 forestry have likewise recorded growth in diameter across
large tracts of the globe \citep{Purves-2008}. The second problem is
that, as we outline further below, because mass production is only one
component of plant growth; theory focussing solely on mass production will
 always be limited.

Effect of traits on shade tolerance also porrly udnerstood. Empirical -- shade

(reich 1992, Niinements 2010, Gower-1993 - both show higher LAI in high LMA / long LL stands)

Ultimately, trait based understanding integrated wiht model of plant growth, also able to cpature patterns of growth through ontogeny. Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}.  In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover. All growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}. Shade tolerance tends to decrease with increasing size \citep{Lusk-2006}, due to increasing costs of mainatining stems \citep{Givnish-1988}

In this paper we provide a mechanistic framework for describing the potential
effect of traits on growth throughout a plant's life-cycle, linking to
commonly used metrics such as height and diameter growth. To help motivate our
study we re-analyse demographic and trait data from an earlier study
 \citep{Wright-2010}, to show 1) that the link between traits and
growth is stronger than previously thought, 2) that the influence of three
prominent traits on growth rate is moderated by plant size, and 3) that the
impact of size on the nature of relationships varies among traits
 \citep[see also]{Ruger-2012}. We then show how these results align with new
set of theoretical predictions. Analysis extends ...
Moreover, we show how traits moderate plant
responses to light environment and also determine shade tolerance, supporting
diverse empirical results \citep{Ruger-2012, Poorter-2006}. By
disentangling the effects of plant size, light environment and traits on
growth rates, we hope to provide a solid theoretical foundation for trait
ecology and thus provide a platform for understanding growth across diverse
species around the world.

\section{Conceptual framework for understanding height, basal area and diameter growth}

To make sense of empirical patterns, we extend a widely-used theoretical model that links growth rate in
seedlings to LMA \citep{Lambers-1992, Wright-2000}, to explicitly include influences of
size, light environment, and other prominent traits. This model

Builds on 2011 paper - highlight what's new from there.

First derive ... generally true.

Equations \ref{eq:dbdt} -\ref{eq:dDdt} shown in Fig. \ref{fig:conceptual}.
\setcounter{equation}{6}

Requires definitions from Table \ref{tab:definitions}.

(check this material covered in table: Here, $M, r$, and $k$ refer to the mass, respiration rate, and
turnover rate of different tissues, denoted by subscripts $l$=leaves,
$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
rate of CO$_2$ per leaf area and $y$ is yield: the fraction of
assimilated carbon fixed in biomass (the remaining fraction being lost
as growth respiration) (see Table \ref{tab:params} for units and
definitions). Photosynthesis is proportional to leaf area,
$A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per area.
The total mass of living tissues $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}.$
)

We begin with a standard model for the amount of biomass available for
growth, ${\rm d}B / {\rm d}t$, given by the difference between income
(total photosynthesis) and losses (respiration and turnover) within the
plant \citep{Makela-1997, Thornley-2000, Falster-2011}:

\begin{equation}\label{eq:dbdt_full}
\underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
  = \underbrace{\strut y}_\textrm{yield}
    \big( \underbrace{\strut A_{\rm l} \, p}_\textrm{photosynthesis} -
     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
\end{equation}

The only potentially controversial element of in eq. \ref{eq:dbdt} \ref{eq:dbdt_full} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}. This assumption is likely to hold true for most woody plants, because the ..... However, the assumption may not hold for some herbs or annuals, where the swicth to reproduction often entails run-down in the vegetative part of the plant (refs XX) \citep{ Thornley-2000}.

To model growth in either height, basal area or stem diameter requires that account not just for mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout. Mathematically, height growth can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Figure \ref{fig:conceptual})

The first term on the right of eq \ref{eq:dhdt},
${\rm d}H / {\rm d}A_{\rm l}$, is the growth in plant height
per unit growth in total leaf area -- accounting for the architectural
strategy of the plant. Some species tend to leaf out more than grow
tall, while other species emphasise vertical
extension \citep{Poorter-2006}.

The second term, ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$,
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures. As such, ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$
can itself be expressed as a sum of construction costs per unit leaf
area:

The third term in eq \ref{eq:dhdt},
${\rm d}M_{\rm a} / {\rm d}P$, gives the fraction of net
biomass production (eq. \ref{eq:dbdt}) that is allocated to growth
rather than reproduction or storage.

In a similar way, basal area ($A_{\rm st}$) increment can be
expressed as the sum of increments in sapwood, bark \& heartwood areas
($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively):
$\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
We now have an equation for basal area growth that contains many of the same
elements as eq. \ref{eq:dhdt}, see eq. \ref{eq:dast}.

Diameter growth is then given by the geometric relationship between stem
diameter ($D$) and $A_{\rm st}$, see eq. \ref{eq:dDdt}

\subsubsection{A functional-balance model for allocation}

To address the first of these challenges, we use a functional balance
model \citep{Yokozawa-1995, Falster-2011} assuming
1) an allometric relation between total leaf area and plant height, and
2) a constant ratio between total leaf area and each of sapwood
cross-sectional area, bark cross-sectional area, and total root mass.
These assumptions, which are well-supported by data (Fig.
\ref{fig:assumptions}), lead immediately to an allometric model for plant
size (Table \ref{tab:allometry}, see Supplementary material for
derivations). Substituting from Table \ref{tab:allometry} into eqs.
\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
for plant growth in relation to size.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  four key assumptions.
\item
  simple linear functions non-critical for results of this paper,
\item
  explore using an example model, but any model with basic size effects
  will display same behaviour

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    details about our model ()
  \item
    show argument hold intuitively
  \end{itemize}
\end{itemize}

\section{Derivation of a simple allometric model of plant function}

Here we describe an allometric model linking the various size dimensions
of a plant required by most ecologically realistic vegetation models
(i.e. =mass of leaves, mass of sapwood, mass of bark, mass of fine
roots) to a plant height. Table \ref{tab:allometry} provides a summary
of the derived model, while Table \ref{tab:params} provdies estimates on
key parameters.

Based on empirically observed allometries (see main text), we assume an
allometric log-log scaling relationship between the accumulated leaf
area of a plant and its height: Eq. \ref{eq:ha}.

Note, scaling relationship reversed from \citep{Falster-2011}.

We follow the model of \citep{Yokozawa-1995} describing the
vertical distribution of leaf area within the crowns of individual
plants. This model can account for a variety of canopy profiles through
a single parameter $\eta$. Setting $\eta=1$ results in a conical
canopy, as seen in many conifers, while higher values, e.g. $\eta=12$
, give a top-weighted canopy profile similar to those seen among
angiosperms. We also assume that each unit of sapwood area supports a fixed area of
leaf (the pipe model, \citep{Shinozaki-1964}), so that the total
canopy area of a plant relates to basal sapwood area. This leads to a solution for the total mass of sapwood in the plant as a function leaf area and height (Eq. \ref{eq:ms2}).

Bark and phloem tissue are modelled using an analogue of the pipe model,
leading to a similar equation as that for sapwood mass (eq.
\ref{eq:ms2}). Cross sectional-area of bark per unit leaf area is
assumed to be a constant fraction $b$ of sapwood area per unit leaf
area such that (Eq. \ref{eq:mb}).

Also consistent with pipe-model assumption, we assume a fixed ratio of
root mass per unit leaf area (Eq. \ref{eq:mr}).

Even though nitrogen and water uptake are not modelled explicitly,
imposing a fixed ratio of root mass to leaf area ensures that
approximate costs of root production are included in calculations of
carbon budget.

\subsubsection{Embedding trait-related trade-offs}

use table with arrows to

By definition, the main effect of increasing $\phi$ is to decrease the
marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
growth rate. However, to fully account for the influence of $\phi$ on
growth we must also account for the decrease in leaf turnover that
results from superior leaf construction \citep{Wright-2004}.
This is achieved by linking the turnover rate of leaf in eq.
\ref{eq:dbdt} to $\phi$. We use an allometric scaling relationship of
the form $k_{\rm l}=\alpha_4 \, \phi^{-\beta_4}$, which has been
shown to describe patterns across diverse
ecosystems \citep{Wright-2004} (Fig. \ref{fS-leaf}).

>>>>

% Enable new column type -- centered with fixed width
% Following http://texblog.org/2008/05/07/fwd-equal-cell-width-right-and-centre-aligned-content/
\usepackage{array}
\newcolumntype{x}[1]{%
>{\raggedleft}p{#1}}%

@@

% Enable cross referencing to external documents
\usepackage{xr}
\externaldocument{ms-supp}
\externaldocument{figure_conceptual}
% Needed to set equation counter, following inclusion of figure 1
\setcounter{equation}{7}

% Input common definitions from external file
\input{ms-common-defs}
\date{}

@@

\keywords{growth strategy, life-history,  ontogeny, size, leaf mass per area, wood density, shade-tolerance, growth rate}

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
% \doublespacing
% \linenumbers

\tableofcontents
% Consider Ecol Letters "Ideas and Perspectives": maximum of 7500 words (main text), and 10 figures, tables or boxes.

\section{Abstract}
We present a new theoretical framework for understanding the effect of diverse functional traits on plant growth.

@@

Functional traits capture core differences in the strategies plants use to generate and invest surplus energy \citep{Wright-2004, Chave-2009, Westoby-2002}. Although most plants have the same basic physiological function and key resource requirements (carbon, nitrogen and water), species differ considerably in rates at which resources are acquired and invested into different tissues. During the last decade, trade-offs related to prominent traits -- such as leaf mass per leaf area, wood density, leaf nitrogen per area, vessel diameter, maximum  height, and seed size -- have been quantified and trait values exist for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}. Increasingly, researchers have been using traits as potential indicators of the growth strategy and ecology of a species.

While the influence of plant traits on elements of plant physiological function is increasingly understood, attempts at using traits to understand demographic rates however have only met with mixed success \citep{Wright-2010, Poorter-2008}. In seedlings, the trait leaf mass per leaf area -- the central element of the leaf economics spectrum \citep{Wright-2004} -- has been tightly related to relative growth rate in plant mass \citep{Lambers-1992, Wright-2000}. Leaf mass per leaf area and its correlate leaf lifespan have also been linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}. These early successes prompted researchers to search for similar relationships in large plants. However, the results showed that in trees leaf mass per leaf area was not correlated with diameter growth rate \citep{Poorter-2008, Wright-2010, Herault-2011}. Meanwhile, other traits such as wood density showed strong relationships to growth in large plants \citep{Wright-2010}, but only weak relationships in seedlings \citep{Castro-1998}. Yet, until recently, most trait growth relationships had been quantified for one or few traits and at a single life- stage.

The picture is now emerging that the effect of traits on plant growth may be  modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}. In a recent meta-analysis of more than  500 correlations recorded from 103 studies, \citet{Gibert-2016} showed that the sign of the correlation between some traits (such as leaf mass per leaf area and maximum height) and growth rate was modified by size, while for other traits (such as wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size. Importantly, the direction of the correlations and their shifts with size, supported predictions made based on understanding of trait-based trade-offs and their interaction with plant function \citep{Gibert-2016}.

The challenge of interpreting diverse empirical results seeking to link traits to growth rate has been all the harder because we currently lack clear expectations on the expected signal. Generating such expectations is one of the primary role for theory \citep{Kokko-2007}. Current empirical results suggest the effect of traits on growth changes with size and possibly also light environment, yet current theory says little about how such relationships might come about. A widely-used model shows that for seedlings, mass-based relative growth rate is linearly and negatively related to leaf mass per leaf area \citep{Lambers-1992, Cornelissen-1996, Wright-2000}. An extension of the model suggests a similar relationship may hold at larger sizes \citep{Enquist-2007}. But as described above, this prediction is not supported by empirical results. Meanwhile, theoretical predictions on how other traits should influence growth are largely absent..

A further problem with existing theory is that the effects of traits are all realised via influences on mass production \citep{Enquist-2007}, whereas the effect of traits such as leaf mass per leaf area and wood density is via allocation to different tissues \citep{Falster-2011, Gibert-2016}. A second issue with theory focussed solely on mass production is that measuring mass production is only practical for small plants that can be easily harvested. For larger plants, growth is more often measured as increment in either diameter or height. For example, the Centre for Tropical Science monitors diameter growth in 49 large permanent plots across forests worldwide \citep{Anderson-2015}; forestry agencies have likewise recorded growth in diameter in 1000's of plots across the globe \citep{Purves-2008, Kunstler-2015}. But as we outline further below, mass production is only one component of plant growth; meaning theory focussing solely on mass production will always be limited.

Ultimately, trait based understanding must b integrated within a more general model of plant growth, also able to capture patterns of growth through ontogeny. Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}.  In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover. All growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}. Shade tolerance tends to decrease with increasing size \citep{Lusk-2006}, due to increasing costs of maintaining stems \citep{Givnish-1988}, and also vary with traits \citep{Lusk-2006,Poorter-2006}. These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are in need of a comprehensive and integrated explanation.

In this paper we show how a recently published mechanistic framework \plant{} \citep{Falster-2016} describing the potential effect of traits on growth throughout a plant's life-cycle can explain diverse phenomena (Table \ref{tab:phenomena}). The model employs the same framework for partitioning growth used by \citet{Gibert-2016} in their meta-analysis, building on and extending several related approaches \citep{Givnish-1988, Yokozawa-1995, Makela-1997, Moorcroft-2001, Falster-2011, King-2011} in which metrics such as height, diameter and mass growth are calculated as an emergent outcomes of several, independent, physiological processes. This approach is quite different to models derived from metabolic scaling theory \citep{Enquist-1999, Enquist-2007}, which attempt to organise everything around a single master ``scaling'' equation for mass growth, or statistically fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014}. By disentangling the effects of plant size, light environment and traits on plant growth, we hope to provide a solid theoretical foundation for trait ecology and thus provide a platform for understanding growth across diverse species around the world.

\section{Framework for understanding the effects of traits on growth}

Our approach builds on the widespread approach used in many vegetation models of explicitly modelling the amounts in different tissues within the plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig. \ref{fig:conceptual}a). Here we consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, with the subscript $i$ indicating: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=live (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h). The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$. A full list of variables, units and definitions is given in Table \ref{tab:params}.

We assume growth is fundamentally driven by energy income and subsequent distribution throughout the plant. Adopting a standard model, the amount of biomass available for growth, ${\rm d}B / {\rm d}t$ is given by the difference between income (total photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000, Falster-2011} (Fig. \ref{fig:conceptual}a, eqn \ref{eq:dbdt}). Photosynthesis is proportional to the average photosynthetic rate per unit leaf area $\bar{p}$ and total leaf area, $A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per leaf area. The constant $y$ in  eqn \ref{eq:dbdt} converts mol CO$_2$ into units of biomass. Although many models include a detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, net biomass production eventually comes down to simple subtraction of income and loss.

The only potentially controversial element of in eqn \ref{eq:dbdt} is that it assumes tissues lost via turnover are replaced before energy is allocated to either new growth (i.e. growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$ or reproduction \citep{Thornley-2000}. This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction often entails run-down in the vegetative part of the plant (refs XX).

\subsection{Decomposing of growth rate into components}

To model growth in either height, leaf area, basal area, standing mass or stem diameter requires that we account not just for net mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout. Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Fig. \ref{fig:conceptual}b, eqns \ref{eq:dbdt}-\ref{eq:dDdt}). Importantly, most of these terms vary intrinsically with size. The insets in Fig. \ref{fig:conceptual}b show size-related patterns for a typical plant, from a specific allocation model derived below. However, it is important to recognise that eqns \ref{eq:dbdt}-\ref{eq:dDdt} are mathematically true, derived using the chain rule of differential calculus, and thus must hold true for any class of models where carbon allocation is important.

The first point to note is that the growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depend on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$. We assume the remaining fraction, $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$, is allocated to reproduction. The term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, presumably at 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants. It is important to note here that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover. So even a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even if the net amount of live mass $M_{\rm a}$ does not increase any further.

The growth rate in total standing mass of the plant, $\frac{{\rm d}M_{{\rm t}}}{{\rm d}t}$ is then sum of sapwood turnover (which is replaced by an equivalent amount) and growth in live mass (eqn \ref{eq:dMtdt}).

The remaining growth rates (eqns \ref{eq:daldt}-\ref{eq:dDdt}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and support tissues of bark, sapwood and roots (eqn \ref{eq:daldmt}). The inverse of this term, $\frac{{\rm d}M_{\rm a} }{{\rm d}A_{\rm l}}$, is the whole plant construction cost per unit leaf area, which can be further decomposed as a sum of tissue-level construction costs per unit of leaf area, with one of these being $\phi = \frac{{\rm d}M_{\rm l} }{{\rm d}A_{\rm l}}$: the trait leaf mass per area.

The rate of height growth (eqn \ref{eq:dhdt}) depends on an additional term, $\frac{{\rm d}H }{{\rm d}A_{\rm l}}$: the growth in plant height per unit growth in leaf area. This variable accounts for the architectural strategy of the plant. Some species tend to leaf out more than grow tall, while other species emphasise vertical extension \citep{Poorter-2006}. Meanwhile, basal area ($A_{\rm st}$) increment can be expressed as the sum of increments in sapwood, bark \& heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$. These in turn are related to ratios of sapwood or bark area per leaf area, and sapwood turnover (see eqn \ref{eq:dast}). Finally, diameter growth is then given by the geometric relationship between stem diameter ($D$) and $A_{\rm st}$ (eqn \ref{eq:dDdt}).

\subsection{A functional-balance model for plant construction}

As noted above, most of eqns in Fig. \ref{fig:conceptual} arise as mathematical truisms. To make predictions from eqns \ref{eq:dbdt}-\ref{eq:dDdt}, however, requires a specific model of plant construction. The model described below can be considered a first-order functional-balance model, in that it seeks to maintain a balance between various elements of plant construction through ontogeny. Table \ref{tab:allometry} provides key equations of the model, while Table \ref{tab:params} provides estimates on key variable and fitted parameters.

Our model is founded on four key assumptions:

\noindent{\bf Assumption  1} -- There is an allometric (log-log) scaling relationship between the accumulated leaf area of a plant and its height. Relationship represented by eqn \ref{eq:ha}. Empirically motivated. Describes architectural strategy of plant.

\noindent{\bf Assumptions 2-4} -- An isometric relationship between the total leaf area and each of sapwood cross-sectional area (2), bark cross-sectional area (3), and total root mass (4). As per the pipe model, \citep{Shinozaki-1964} we assume that each unit of sapwood area supports a fixed area of leaf, so that the total canopy area of a plant relates to basal sapwood area eqn \ref{eq:as}. This leads to a solution for the total mass of sapwood in the plant as a function leaf area and height (eqn \ref{eq:ms2}). Bark and phloem tissue are modelled using an analogue of the pipe model (eqn \ref{eq:as}), leading to a similar equation as that for sapwood mass (eqn \ref{eq:msb}). Also consistent with pipe-model assumption, we assume a fixed ratio of root mass per unit leaf area (eqn \ref{eq:mr}).

\noindent{\bf Assumption 5} -- Vertical distribution of leaf within the canopy of a plant remains relatively constant through ontogeny. We follow the model of \citep{Yokozawa-1995} describing the vertical distribution of leaf area within the crowns of individual plants. This model can account for a variety of canopy profiles through a single parameter $\eta$. Setting $\eta=1$ results in a conical canopy, as seen in many conifers, while higher values, e.g. $\eta=12$ , give a top-weighted canopy profile similar to those seen among angiosperms.

Assuming lead immediately to an allometric model for plant size (Table \ref{tab:allometry}, see Supplementary material for derivations).

To first approximation, the assumptions, which are well-supported by data (Fig. \ref{fig:assumptions}).

Simple linear functions non-critical for results of this paper, explore using an example model, but any model with basic size effects will display same behaviour.

Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model for plant growth in relation to size.

\subsection{Trait-based trade-offs}

\subsection{Parameters}

\citep{Falster-2015b}: BAAD
\citep{Falster-2016}: \plant

\section{Analysis}

This new growth model captures the intrinsically size-dependent nature of plant growth, recovering well-known empirical patterns \citep{Sillett-2010, King-2011}. Height growth shows a hump-shaped pattern with size, first increasing then decreasing. This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with size (Fig. \ref{fig:conceptual}), including a strong decline in the fraction of plant that is leaf declines with increasing size (Fig. \ref{fig:conceptual}), and increasing reproductive allocation. In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover. Diameter growth shows a weakly hump shaped curve \citep{Herault-2011}, tapering off slightly at larger sizes, in part because of allometric effects (eqn \ref{eq:dDdt}, but mainly because of increased reproductive allocation in older trees (Fig?). All growth measures decrease sharply with size when expressed as relative growth rates \citep{Iida-2014}.

Distinguish 3 types of trait.

Growth traits

For growth traits can derive rule:
\begin{equation}\label{eq:dg4}
\frac{\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}} {\frac{ {\rm d}B} {{\rm d}t}} < \frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
\end{equation}
In words, there will be a negative relationship between trait x and growth if the relative change in production due to change in traits is less than the relative change in the total construction cost for a unit of leaf area due to change in trait. Crucially, key elements in eqn \ref{eq:dg3} are modified by height.

We can also solve for trait values $x$ maximising growth rate, i.e. the value giving $\partial G /\partial x = 0$:
\begin{equation}\label{eq:max}
\frac{\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} } {\frac{ {\rm d}B} {{\rm d}t}} = \frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
\end{equation}

\subsubsection{Seed size, $M_{\rm t,0}$}

Seed size: increases starting size

\subsubsection{Height at maturation, $H_{\rm m}$}

Height at maturation: ($h_m$) moderates growth by adjusting the amount of energy invested in growth, i.e.~the term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.

\citet{Wenk-2015}

\subsubsection{Leaf mass per leaf area, $\phi$}

By definition, the main effect of increasing $\phi$ is to decrease the marginal leaf deployment per mass (eqn \ref{eq:daldmt}) and thus reduce growth rate. However, to fully account for the influence of $\phi$ on growth we must also account for the decrease in leaf turnover that results from superior leaf construction \citep{Wright-2004}. This is achieved by linking the turnover rate of leaf in eqn \ref{eq:dbdt} to $\phi$. We use an allometric scaling relationship of the form $k_{\rm l}=\alpha_4 \, \phi^{-\beta_4}$, which has been shown to describe patterns across diverse ecosystems \citep{Wright-2004} (Fig. \ref{fS-leaf}).

Unlike eqn \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and \ref{eq:dDdt} predict a relationship between $\phi$ and growth that changes with plant size ( \ref{fig:lmA_growth_size}). Decreasing $\phi$ has two impacts on growth rate. First, lower $\phi$ increases marginal leaf deployment (${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$ by economising on construction costs. Second, lower $\phi$ decreases net production (${\rm d}P / {\rm d}t$), due to increased leaf turnover. Whether lower $\phi$ increases growth thus depends on the relative magnitude of these two effects. When plants are small the effect on leaf deployment rate is bigger and so decreasing $\phi$ increases growth rate. When plants are large, the influence of $\phi$ on leaf deployment rate is diminished, because the costs of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) are larger (Fig. \ref{fig:conceptual}). The net result is that at larger sizes, low $\phi$ is no longer advantageous for growth (Fig. \ref{fig:lmA_growth_size}d).

If the trait of interest is leaf-construction cost, i.e. $x=\phi$, we can expand eqn \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{{\rm d}M_\textrm{t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$. Also, from eqn \ref{eq:dbdt} we obtain,
\begin{equation}\label{eq:dbdt2}
\frac{{\rm d}B}{{\rm d}t} = c_2 - A_{\rm l} A_4 \phi ^{1-B_4}
\end{equation}
where $c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_\textrm{i=b,s,r}{M_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{M_\textrm{i} \, hmax\textrm{i}})$is independent of $\phi$, and thus
\begin{equation}\label{eq:dbdt3}
\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial \phi}  =
(B_4-1) A_{\rm l} A_4\phi ^{-B_4}.
\end{equation}
eqn \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ {\rm d}B} {{\rm d}t}$ is independent of $\phi$. In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ \textrm{d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
